### 2020-02-21 manuscript/ERL_global_C_review.Rmd
format cleanup

@@

*Background.* The fate of Earth’s climate closely linked to forests, which strongly influence atmospheric carbon dioxide (CO2) and climate through their influential role in the global carbon (C) cycle. Synthetic understanding of global forest C cycles is needed to constrain model estimates of forest feedbacks to climate change and to more accurately quantify the influence of land use decisions on climate.

@@

Forest ecosystems globally influence climate through their critical role in the global carbon (C) cycle (Fig. 1). Their annual gross CO2 sequestration (gross primary productivity, GPP) is estimated at 59 Gt C yr-1 (Beer et al 2010), or 6.3 times average annual fossil fuel emissions from 2007-2016 (9.4 ± 0.5 Gt C yr-1; Le Quéré et al 2017). A small portion of global terrestrial GPP is retained in ecosystems (mainly forests), resulting in a C sink that averaged 3.0 ± 0.8 Gt C yr-1 from 2007-2016, offsetting 32% of anthropogenic fossil fuel emissions (Le Quéré et al 2017). The remaining ~98% of global GPP is counterbalanced by ecosystem respiration (Reco) or wildfire. Perturbation to the global GPP- Reco balance can substantially influence the growth rate of atmospheric CO2; for example, the 2015-2016 El Niño, which brought historically high temperatures and low precipitation to the tropics, released an extra 2.5 ± 0.3 Gt C to the atmosphere, resulting in the largest recorded atmospheric CO2 growth rate (Le Quéré et al 2016, Liu et al 2017). In addition, forests contain substantial C stocks that, when disturbed, release significant amounts of CO2 to the atmosphere. Although they cover only ~30% of the land surface, forests contain an estimated 92% of terrestrial biomass (Pan et al 2013) and 45% of terrestrial C (biomass and soils; Bonan 2008). Globally, gross tropical deforestation averaged 2.8 Gt C yr-1 from 2000-2007, but ~40% of this was offset by forest regrowth, resulting in a net source of ~1.1 Gt C yr-1 from tropical land use change (Pan et al 2011). This, coupled with minimal net deforestation in the extratropics and net uptake by intact forests (Pan et al 2011, Schimel et al 2015), resulted in a total gain in forest C of ~1.2-1.7 Gt C yr-1 (Le Quéré et al 2017, Schimel et al 2015), thereby substantially slowing the rate of increase of atmospheric CO2.

![**Figure 1 | The role of forests in the global carbon cycle. Values apply to the period 2000-2018.** Sources are as follows: biomass- (Pan et al 2013, Baccini et al 2012) IPCC 2000; intact forest sink, net forest sink, tropical deforestation and regrowth-(Pan et al 2011); terrestrial sink- (Le Quéré et al 2017); GPP- (Beer et al 2010); respiration+fire: calculated here; fossil fuel emissions- (Le Quéré et al 2017); atmospheric CO2 (2018 value from https://scripps.ucsd.edu/programs/keelingcurve/). **see issue 19: https://github.com/forc-db/ERL-review/issues/19** ](tables_figures/Global_C_cycle/Global C cycle diagram.png)

Given their vital role in regulating atmospheric CO2, the future of Earth’s forests will strongly influence the course of climate change. If forests globally respond to the suite of contemporary and future global change drivers—including elevated CO2, climate change, and atmospheric deposition—with increased productivity and net biomass increases, they could continue to act as a significant buffer against anthropogenic emissions. In contrast, if factors such as higher temperatures and drought stress lead to net C losses, the C sink of forests could be lost and even reversed. Global coupled climate models vary substantially in predictions regarding the future of the global forest C sink, with uncertainty in end-of-century terrestrial NEP exceeding current annual anthropogenic CO2 emissions (Friedlingstein et al 2006, Cavaleri et al 2015). Anthropogenic land use decisions will also strongly influence the future course of atmospheric CO2. In the absence of measures to reduce deforestation, this will continue to be a significant component of anthropogenic CO2 emissions. On the other hand, forest conservation, reforestation, and forest management hold strong potential to help stabilize and eventually reduce total anthropogenic CO2 emissions (estimated total of 4.4 Gt C yr-1; Griscom et al 2017; *Cook-Patton et al.*; **other refs?**) and thereby contribute towards the goals of the Paris Climate Agreement (UNFCCC 2015, Houghton et al 2015; **REFS**). Indeed, forest-based climate mitigation was a key component of Paris Climate Agreement commitments, totaling approximately one-fourth of nations’ planned emissions reductions (Grassi et al 2017; IPCC1.5). Thus, forest will strongly influence the future course of climate change both through their responses to global change and through forest-related anthropogenic land use decisions.

While forests will play a critical role in the future course of climate change, there remain significant uncertainties as to current and future C cycling the world’s forests. Even as the climate changes and forest age distributions shift, there are important gaps in our fundamental understanding of how C cycling varies with climate and stand age. A relative dearth of data on C cycling in secondary forests, particularly in the tropics (Anderson-Teixeira et al 2016), is problematic in that almost 2/3 of the world’s forests were secondary as of 2010 (FAO 2010), implying an under-filled need to characterize age-related trends in forest C cycling. Recent studies attempting to quantify the role of forests in the global C cycle (e.g., Pan et al 2011) have been hampered by insufficient data on regrowth rates of tropical forests worldwide, C in non-living pools (i.e., dead wood, litter, soil), and source/sink status of forests in some regions. Moreover, as discussed above, model representation of forest responses to global change remain highly uncertain (Friedlingstein et al 2006, Cavaleri et al 2015), with efforts to reduce model uncertainty in part limited by ready availability of appropriate benchmark data (Clark et al 2017, Luo et al 2012). To address the need for global-scale analyses of forest C cycling, we have developed an open-access Global Forest Carbon database, ForC (https://github.com/forc-db/ForC; Anderson-Teixeira et al 2016, 2018). ForC contains data on forest ecosystem C stocks and annual fluxes (>50 variables) and associated data required for interpretation (e.g., stand history, measurement methods) amalgamated from numerous previous data compilations and directly from original publications. ForC currently contains **#** records from **#** plots and **#** distinct geographic areas representing all forested biogeographic and climate zones.

@@

1. How do C cycling vary across the worlds major forest biomes (i.e., tropical, temperate, boreal)?

@@

While components of these questions have been previously addressed (e.g., REFS(Luyssaert et al 2007, Anderson et al 2006, Anderson-Teixeira et al 2013), our analysis represents by far the most comprehensive analysis of C cycling in global forests, and thereby stands to serve as a foundation for improved understanding of global forest C cycling.

@@

This review synthesizes data from the ForC database (https://github.com/forc-db/ForC; Anderson-Teixeira et al 2016, 2018). ForC amalgamates numerous intermediary data sets (e.g., REFS) and original studies. Original publications were referenced to check values and obtain information not contained in intermediary data sets, although this process has not been completed for all records. The database was developed with goals of understanding how C cycling in forests varies across broad geographic scales and as a function of stand age. As such, there has been a focus on incorporating data from regrowth forests (e.g., Anderson et al 2006, Martin et al 2013, Bonner et al 2013) and obtaining stand age data when possible (83% of records in v.2.0; Anderson-Teixeira et al 2018). Particular attention was given to developing the database for tropical forests (Anderson-Teixeira et al 2016), yet these represented only approximately one-third of records in ForC v.2.0 (Anderson-Teixeira et al 2018). Since publication of ForC v.2.0, we added the following data to ForC: the Global Database of Soil Respiration Database (Bond-Lamberty and Thomson 2010; *hopefully Cook-Patton*; see [GitHub list](https://github.com/forc-db/ForC/blob/master/database_management_records/ForC_data_additions_log.csv)) …. We note that there remains a significant amount of relevant data that is not yet included in ForC, particularly biomass data from national forest inventories (e.g.,: REFS). The database version used for this analysis has been tagged as a new release on Github (XX) and assigned a DOI through Zenodo (DOI: XX).

@@

We drew upon records for *#* annual flux and *#* C stock variables (Table 1). For this analysis, we combined some of ForC’s specific variables (e.g., multiple variables for net primary productivity including various components) into more broadly defined variables (Table 1, [this table](https://github.com/forc-db/ForC/blob/master/figures/C_cycle_diagrams/ForC_variables_mapping_for_C_cycle_diagrams.csv))). Although ForC contains information that may be used to standardize or control for methodological differences (e.g., area sampled, min stem diameter sampled, allometric equations applied; Anderson-Teixeira et al 2018), for this analysis we included all relevant data in ForC. Throughout ForC, for all measurements drawing from tree census data (e.g., biomass, productivity), the minimum stem diameter sampled was $\le$ 10cm. All records were measured directly or derived from field measurements (as opposed to modeled).

@@

Analyses drew from ForC-simplified (https://github.com/forc-db/ForC/blob/master/ForC_simplified), which is a rearrangement of ForC intended to facilitate analyses. In generating ForC-simplified, all measurements originally expressed in units of dry organic matter ($OM$) were converted to units of C using the IPCC default of $C = 0.47 * OM$ (IPCC 2006). Duplicate or otherwise conflicting records (indicated in conflicts field of MEASUREMENTS table) were reconciled as follows. Replicate measurements (i.e., replicates from within a single study) were averaged. Records that subsumed others— *i.e.*, the time period included that of $\ge$ 2 other records or dates were unknown and therefore conflicted with $\ge$ 2 other records—were removed. For each group of duplicate records—i.e., measurements of the same variable in the same plot at the same time—one record was assigned precedence (recorded in D.precedence field). When measurement periods overlapped or were not specified, precedence was given first to records representing longer measurement periods (*i.e.*, end.date - start.date) and then to more recently published values. We manually reviewed duplicates that differed only in methodology, assigning precedence to the record employing a more comprehensive approach (e.g., inclusion of understory, lianas, or bamboo as opposed to just trees) or using a favored methodology.

Records were filtered to remove plots that had undergone significant anthropogenic management or major disturbance since the most recent stand establishment (*i.e.*, that reflected by stand.age). Specifically, we removed all plots flagged as managed in ForC-simplified (managed field). This included plots with any record of managements manipulating CO2, temperature, hydrology, nutrients, or biota, as well as any plots whose site or plot name contained the terms “plantation”, "planted", “managed”, “irrigated”, or “fertilized”.  Plots flagged as disturbed in ForC-simplified included stands that had undergone anthropogenic thinning or partial harvest (“Cut” or “Harvest” codes) unless this was very minor (percent.mortality= “minor”). We retained sites that were grazed or had undergone low severity natural disturbances (<10% mortality) including droughts, major storms, fires, and floods. We also removed all plots for which no stand history information had been retrieved.

Data were analyzed to produce basic summaries of C cycle patterns across biomes and stand ages following an approach similar to that of Anderson-Teixeira et al (2016). For mature forests, to obtain the values reported in the C cycle schematics, we first averaged any repeated measurements within a plot, weighting flux measurements according to the length of measurement periods (i.e., end.date - start.date). Values were then averaged across plots clustered within 25 km of one another (geographic.area field of SITES table, sensu Anderson-Teixeira et al 2018), weighting by area.sampled (MEASUREMENTS table) or plot.area (PLOTS table) if available for all records. This step was taken to avoid pseudo-replication and to combine any records from sites with more than one name in ForC. Finally, we computed statistics with geographic.area as the unit of replication. To compare across biomes, [**Valentine, please describe**]. There were enough data to run this analysis for all focal variables but XX.

@@

Average C cycles for tropical broadleaf, temperate broadleaf, temperate conifer, and boreal forests $\ge$ 100 years old and with no known major disturbance or significant anthropogenic management are presented in Figures 2-5, and statistics for each biome type are also summarized at: [GitHub URL]. Of the # flux and # stock variables mapped in these diagrams, ForC contained estimates from $\ge$ 7 distinct geographic areas for *#* fluxes and *#* stocks in tropical broadleaf forests, # fluxes and # stocks in temperate broadleaf forests, #fluxes and # stocks in temperate conifer forests, and fluxes and # stocks in boreal forests. For variables with records from $\ge$ 7 distinct geographic areas, these ensemble C budgets were generally consistent. That is, component fluxes and stocks summed to within 1 std of more inclusive fluxes in all but one instance (in temperate conifer forests, $aboveground woody biomass + foliage biomass > aboveground biomass + 1std$; Fig. 4). **This is based on my quick scan of the figures. Need to add a formal automated test: https://github.com/forc-db/ERL-review/issues/16 **

@@

There were less distinct trends in C stocks across biomes (Fig. 7). Biome means for live aboveground and total biomass followed the same general trend as the major fluxes, with biomass of tropical= temperate broadleaf $\ge$ temperate conifer> boreal forests (Fig. 7a). However, the relative differences in these means were much smaller than for the major C fluxes, and there was significant variation within biomes. Maximum aboveground biomass values followed a very different trend than the means: temperate broadleaf> temperate conifer>boreal>tropical. There were some statistically significant biome differences in less frequently sampled C stocks (e.g., woody biomass, foliage biomass, deadwood), but given high within-biome variability in C stocks and relatively low sample sizes, these were likely attributable to sampling biases and methodological differences than to true differences across biomes.

@@

(**check paragraph with latest data**) ForC contained 14 flux variables with sufficient data for cross-biome analyses of age trends in regrowth forests (see Methods) (Fig. 6-7 and **S#- SI figures including plots for all variables**). Of these, 9 increased significantly with log10[stand.age]: GPP, NPP, ANPP, ANPP_foliage, ANPP_woody, ANPP_woody_stem, BNPP, BNPP_root_fine, Reco, and net C sequestration (-NEE). The remaining five--ANPP_woody_branch, BNPP_root_coarse, soil heterotrophic respiration, and total soil respiration–displayed no significant relationship to stand age, although all displayed a positive trend. In terms of C stocks, 10 variables had sufficient data to test for age trends. Six of these—total biomass, aboveground biomass, aboveground woody biomass, foliage biomass, root biomass, and coarse root biomass—increased significantly with log10[stand.age]. The remaining four displayed non-significant positive trends:  fine root biomass, total dead wood, standing dead wood, and organic layer.

>>>>

                   \usepackage{setspace}
                  \onehalfspacing

@@

*Background.* The fate of Earth’s climate closely linked to forests, which strongly influence atmospheric carbon dioxide (CO~2~) and climate through their influential role in the global carbon (C) cycle. Synthetic understanding of global forest C cycles is needed to constrain model estimates of forest feedbacks to climate change and to more accurately quantify the influence of land use decisions on climate.

@@

Forest ecosystems globally influence climate through their critical role in the global carbon (C) cycle (Fig. 1). Their annual gross CO~2~ sequestration (gross primary productivity, $GPP$) is estimated at 59 Gt C yr-1 (Beer et al 2010), or 6.3 times average annual fossil fuel emissions from 2007-2016 (9.4 ± 0.5 Gt C yr-1; Le Quéré et al 2017). A small portion of global terrestrial $GPP$ is retained in ecosystems (mainly forests), resulting in a C sink that averaged 3.0 ± 0.8 Gt C yr-1 from 2007-2016, offsetting 32% of anthropogenic fossil fuel emissions (Le Quéré et al 2017). The remaining ~98% of global $GPP$ is counterbalanced by ecosystem respiration (Reco) or wildfire. Perturbation to the global $GPP$- Reco balance can substantially influence the growth rate of atmospheric CO~2~; for example, the 2015-2016 El Niño, which brought historically high temperatures and low precipitation to the tropics, released an extra 2.5 ± 0.3 Gt C to the atmosphere, resulting in the largest recorded atmospheric CO~2~ growth rate (Le Quéré et al 2016, Liu et al 2017). In addition, forests contain substantial C stocks that, when disturbed, release significant amounts of CO~2~ to the atmosphere. Although they cover only ~30% of the land surface, forests contain an estimated 92% of terrestrial biomass (Pan et al 2013) and 45% of terrestrial C (biomass and soils; Bonan 2008). Globally, gross tropical deforestation averaged 2.8 Gt C yr-1 from 2000-2007, but ~40% of this was offset by forest regrowth, resulting in a net source of ~1.1 Gt C yr-1 from tropical land use change (Pan et al 2011). This, coupled with minimal net deforestation in the extratropics and net uptake by intact forests (Pan et al 2011, Schimel et al 2015), resulted in a total gain in forest C of ~1.2-1.7 Gt C yr-1 (Le Quéré et al 2017, Schimel et al 2015), thereby substantially slowing the rate of increase of atmospheric CO~2~.

![**Figure 1 | The role of forests in the global carbon cycle. Values apply to the period 2000-2018.** Sources are as follows: biomass- (Pan et al 2013, Baccini et al 2012) IPCC 2000; intact forest sink, net forest sink, tropical deforestation and regrowth-(Pan et al 2011); terrestrial sink- (Le Quéré et al 2017); $GPP$- (Beer et al 2010); respiration+fire: calculated here; fossil fuel emissions- (Le Quéré et al 2017); atmospheric CO~2~ (2018 value from https://scripps.ucsd.edu/programs/keelingcurve/). **see issue 19: https://github.com/forc-db/ERL-review/issues/19** ](tables_figures/Global_C_cycle/Global C cycle diagram.png)

Given their vital role in regulating atmospheric CO~2~, the future of Earth’s forests will strongly influence the course of climate change. If forests globally respond to the suite of contemporary and future global change drivers—including elevated CO~2~, climate change, and atmospheric deposition—with increased productivity and net biomass increases, they could continue to act as a significant buffer against anthropogenic emissions. In contrast, if factors such as higher temperatures and drought stress lead to net C losses, the C sink of forests could be lost and even reversed. Global coupled climate models vary substantially in predictions regarding the future of the global forest C sink, with uncertainty in end-of-century terrestrial NEP exceeding current annual anthropogenic CO~2~ emissions (Friedlingstein et al 2006, Cavaleri et al 2015). Anthropogenic land use decisions will also strongly influence the future course of atmospheric CO~2~. In the absence of measures to reduce deforestation, this will continue to be a significant component of anthropogenic CO~2~ emissions. On the other hand, forest conservation, reforestation, and forest management hold strong potential to help stabilize and eventually reduce total anthropogenic CO~2~ emissions (estimated total of 4.4 Gt C yr-1; Griscom et al 2017; *Cook-Patton et al.*; **other refs?**) and thereby contribute towards the goals of the Paris Climate Agreement (UNFCCC 2015, Houghton et al 2015; **REFS**). Indeed, forest-based climate mitigation was a key component of Paris Climate Agreement commitments, totaling approximately one-fourth of nations’ planned emissions reductions (Grassi et al 2017; IPCC1.5). Thus, forest will strongly influence the future course of climate change both through their responses to global change and through forest-related anthropogenic land use decisions.

While forests will play a critical role in the future course of climate change, there remain significant uncertainties as to current and future C cycling the world’s forests. Even as the climate changes and forest age distributions shift, there are important gaps in our fundamental understanding of how C cycling varies with climate and stand age. A relative dearth of data on C cycling in secondary forests, particularly in the tropics (Anderson-Teixeira et al 2016), is problematic in that almost 2/3 of the world’s forests were secondary as of 2010 (FAO 2010), implying an under-filled need to characterize age-related trends in forest C cycling. Recent studies attempting to quantify the role of forests in the global C cycle (*e.g.*, Pan et al 2011) have been hampered by insufficient data on regrowth rates of tropical forests worldwide, C in non-living pools (*i.e.*, dead wood, litter, soil), and source/sink status of forests in some regions. Moreover, as discussed above, model representation of forest responses to global change remain highly uncertain (Friedlingstein et al 2006, Cavaleri et al 2015), with efforts to reduce model uncertainty in part limited by ready availability of appropriate benchmark data (Clark et al 2017, Luo et al 2012). To address the need for global-scale analyses of forest C cycling, we have developed an open-access Global Forest Carbon database, ForC (https://github.com/forc-db/ForC; Anderson-Teixeira et al 2016, 2018). ForC contains data on forest ecosystem C stocks and annual fluxes (>50 variables) and associated data required for interpretation (*e.g.*, stand history, measurement methods) amalgamated from numerous previous data compilations and directly from original publications. ForC currently contains **#** records from **#** plots and **#** distinct geographic areas representing all forested biogeographic and climate zones.

@@

1. How do C cycling vary across the worlds major forest biomes (*i.e.*, tropical, temperate, boreal)?

@@

While components of these questions have been previously addressed (*e.g.*, REFS(Luyssaert et al 2007, Anderson et al 2006, Anderson-Teixeira et al 2013), our analysis represents by far the most comprehensive analysis of C cycling in global forests, and thereby stands to serve as a foundation for improved understanding of global forest C cycling.

@@

This review synthesizes data from the ForC database (https://github.com/forc-db/ForC; Anderson-Teixeira et al 2016, 2018). ForC amalgamates numerous intermediary data sets (*e.g.*, REFS) and original studies. Original publications were referenced to check values and obtain information not contained in intermediary data sets, although this process has not been completed for all records. The database was developed with goals of understanding how C cycling in forests varies across broad geographic scales and as a function of stand age. As such, there has been a focus on incorporating data from regrowth forests (*e.g.*, Anderson et al 2006, Martin et al 2013, Bonner et al 2013) and obtaining stand age data when possible (83% of records in v.2.0; Anderson-Teixeira et al 2018). Particular attention was given to developing the database for tropical forests (Anderson-Teixeira et al 2016), yet these represented only approximately one-third of records in ForC v.2.0 (Anderson-Teixeira et al 2018). Since publication of ForC v.2.0, we added the following data to ForC: the Global Database of Soil Respiration Database (Bond-Lamberty and Thomson 2010; *hopefully Cook-Patton*; see [GitHub list](https://github.com/forc-db/ForC/blob/master/database_management_records/ForC_data_additions_log.csv)) …. We note that there remains a significant amount of relevant data that is not yet included in ForC, particularly biomass data from national forest inventories (*e.g.*,: REFS). The database version used for this analysis has been tagged as a new release on Github (XX) and assigned a DOI through Zenodo (DOI: XX).

@@

We drew upon records for *#* annual flux and *#* C stock variables (Table 1). For this analysis, we combined some of ForC’s specific variables (*e.g.*, multiple variables for net primary productivity including various components) into more broadly defined variables (Table 1, [this table](https://github.com/forc-db/ForC/blob/master/figures/C_cycle_diagrams/ForC_variables_mapping_for_C_cycle_diagrams.csv))). Although ForC contains information that may be used to standardize or control for methodological differences (*e.g.*, area sampled, min stem diameter sampled, allometric equations applied; Anderson-Teixeira et al 2018), for this analysis we included all relevant data in ForC. Throughout ForC, for all measurements drawing from tree census data (*e.g.*, biomass, productivity), the minimum stem diameter sampled was $\le$ 10cm. All records were measured directly or derived from field measurements (as opposed to modeled).

@@

Analyses drew from ForC-simplified (https://github.com/forc-db/ForC/blob/master/ForC_simplified), which is a rearrangement of ForC intended to facilitate analyses. In generating ForC-simplified, all measurements originally expressed in units of dry organic matter ($OM$) were converted to units of C using the IPCC default of $C = 0.47 * OM$ (IPCC 2006). Duplicate or otherwise conflicting records (indicated in conflicts field of MEASUREMENTS table) were reconciled as follows. Replicate measurements (*i.e.*, replicates from within a single study) were averaged. Records that subsumed others— **i.e.**, the time period included that of $\ge$ 2 other records or dates were unknown and therefore conflicted with $\ge$ 2 other records—were removed. For each group of duplicate records—*i.e.*, measurements of the same variable in the same plot at the same time—one record was assigned precedence (recorded in D.precedence field). When measurement periods overlapped or were not specified, precedence was given first to records representing longer measurement periods (*i.e.*, end.date - start.date) and then to more recently published values. We manually reviewed duplicates that differed only in methodology, assigning precedence to the record employing a more comprehensive approach (*e.g.*, inclusion of understory, lianas, or bamboo as opposed to just trees) or using a favored methodology.

Records were filtered to remove plots that had undergone significant anthropogenic management or major disturbance since the most recent stand establishment (*i.e.*, that reflected by stand.age). Specifically, we removed all plots flagged as managed in ForC-simplified (managed field). This included plots with any record of managements manipulating CO~2~, temperature, hydrology, nutrients, or biota, as well as any plots whose site or plot name contained the terms “plantation”, "planted", “managed”, “irrigated”, or “fertilized”.  Plots flagged as disturbed in ForC-simplified included stands that had undergone anthropogenic thinning or partial harvest (“Cut” or “Harvest” codes) unless this was very minor (percent.mortality= “minor”). We retained sites that were grazed or had undergone low severity natural disturbances (<10% mortality) including droughts, major storms, fires, and floods. We also removed all plots for which no stand history information had been retrieved.

Data were analyzed to produce basic summaries of C cycle patterns across biomes and stand ages following an approach similar to that of Anderson-Teixeira et al (2016). For mature forests, to obtain the values reported in the C cycle schematics, we first averaged any repeated measurements within a plot, weighting flux measurements according to the length of measurement periods (*i.e.*, end.date - start.date). Values were then averaged across plots clustered within 25 km of one another (geographic.area field of SITES table, sensu Anderson-Teixeira et al 2018), weighting by area.sampled (MEASUREMENTS table) or plot.area (PLOTS table) if available for all records. This step was taken to avoid pseudo-replication and to combine any records from sites with more than one name in ForC. Finally, we computed statistics with geographic.area as the unit of replication. To compare across biomes, [**Valentine, please describe**]. There were enough data to run this analysis for all focal variables but XX.

@@

Average C cycles for tropical broadleaf, temperate broadleaf, temperate conifer, and boreal forests $\ge$ 100 years old and with no known major disturbance or significant anthropogenic management are presented in Figures 2-5, and statistics for each biome type are also summarized at: [GitHub URL]. Of the # flux and # stock variables mapped in these diagrams, ForC contained estimates from $\ge$ 7 distinct geographic areas for *#* fluxes and *#* stocks in tropical broadleaf forests, # fluxes and # stocks in temperate broadleaf forests, #fluxes and # stocks in temperate conifer forests, and fluxes and # stocks in boreal forests. For variables with records from $\ge$ 7 distinct geographic areas, these ensemble C budgets were generally consistent. That is, component fluxes and stocks summed to within 1 std of more inclusive fluxes in all but one instance (in temperate conifer forests, $aboveground woody biomass + foliage biomass > aboveground biomass + 1std$; Fig. 4). **update this: https://github.com/forc-db/ERL-review/issues/16 **

@@

There were less distinct trends in C stocks across biomes (Fig. 7). Biome means for live aboveground and total biomass followed the same general trend as the major fluxes, with biomass of tropical= temperate broadleaf $\ge$ temperate conifer> boreal forests (Fig. 7a). However, the relative differences in these means were much smaller than for the major C fluxes, and there was significant variation within biomes. Maximum aboveground biomass values followed a very different trend than the means: temperate broadleaf> temperate conifer>boreal>tropical. There were some statistically significant biome differences in less frequently sampled C stocks (*e.g.*, woody biomass, foliage biomass, deadwood), but given high within-biome variability in C stocks and relatively low sample sizes, these were likely attributable to sampling biases and methodological differences than to true differences across biomes.

@@

(**check paragraph with latest data**) ForC contained 14 flux variables with sufficient data for cross-biome analyses of age trends in regrowth forests (see Methods) (Fig. 6-7 and **S#- SI figures including plots for all variables**). Of these, 9 increased significantly with log10[stand.age]: $GPP$, $NPP$, $ANPP$, $ANPP_{foliage}$, $ANPP_{woody}$, $ANPP_{woody-stem}$, $BNPP$, $BNPP_{root-fine}$, $R_{eco}$, and net C sequestration ($-NEE$). The remaining five--$ANPP_{woody-branch}$, $BNPP_{root-coarse}$, $R_{soil-het}$, and $R_{soil-het}$-–displayed no significant relationship to stand age, although all displayed a positive trend. In terms of C stocks, 10 variables had sufficient data to test for age trends. Six of these—total biomass, aboveground biomass, aboveground woody biomass, foliage biomass, root biomass, and coarse root biomass—increased significantly with $log10[stand.age]$. The remaining four displayed non-significant positive trends:  fine root biomass, total dead wood, standing dead wood, and organic layer.
